---
title: "Hands-on_Ex04"
author: "LIU YAN"
editor: visual
---
## Install and launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. if they are , then they will be lauched into R.

```{r}
#| warning: false
pacman::p_load(readxl,gifski,gapminder,plotly,gganimate, tidyverse) 
```

## Importing the data

```{r}
#| warning: false
col<-c("Country", "Continent")
globalPop<-read_xls("data/GlobalPopulation.xls",sheet="Data")%>%
mutate_each_(funs(factor(.)), col) %>%
mutate(Year = as.integer(Year))
```
## 4.3 Building a static population bubble plot
```{r}
#| warning: false
ggplot(globalPop, aes(x = Old, y = Young, 
                      size = Population, 
                      colour = Country)) +
  geom_point(alpha = 0.7, 
             show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  labs(title = 'Year: {frame_time}', 
       x = '% Aged', 
       y = '% Young') 
```
## 4.3.2 Building the animated bubble plot
```{r}
#| warning: false
ggplot(globalPop, aes(x = Old, y = Young, 
                      size = Population, 
                      colour = Country)) +
  geom_point(alpha = 0.7, 
             show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  labs(title = 'Year: {frame_time}', 
       x = '% Aged', 
       y = '% Young') +
  transition_time(Year)+
  ease_aes('linear')
```
## 4.4 Animated Data Visualisation: plotly
## 4.4.1 Building an animated bubble plot: ggplotly() method
```{r}


```
## 4.4.2 Building an animated bubble plot: plot_ly() method
```{r}
#| warning: false
bp <- globalPop %>%
  plot_ly(x = ~Old, 
          y = ~Young, 
          size = ~Population, 
          color = ~Continent, 
          frame = ~Year, 
          text = ~Country, 
          hoverinfo = "text",
          type = 'scatter',
          mode = 'markers'
          )
bp
```
