---
title: "Hands-on_Ex03"
author: "LIU YAN"
editor: visual
---
## Install and launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. if they are , then they will be lauched into R.

```{r}
pacman::p_load(ggiraph, plotly, 
               patchwork, DT, tidyverse) 
```

## Importing the data

```{r}
exam_data<-read_csv("data/Exam_data.csv")
```
3.4.1 Tooltip effect with tooltip aesthetic
```{r}
p<-ggplot(data=exam_data,
          aes(x=MATHS))+
  geom_dotplot_interactive(
    aes(tooltip=ID),
    stackgroups=TRUE,
    binwidth=1,
    method="histodot")+
  scale_y_continuous(NULL,breaks = NULL
  )
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)
  
```
3.4.2 Displaying multiple information on tooltip
```{r}
exam_data$tooltip<-c(paste0("Name= ",exam_data$ID,"\n Class = ", exam_data$CLASS))

p<-ggplot(data=exam_data,
          aes(x=MATHS))+
  geom_dotplot_interactive(
    aes(tooltip=exam_data$tooltip),stackgroups=TRUE, binwidth=1,method="histodot")+scale_y_continuous(NULL,breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 8,
  height_svg = 8*0.618
)
  
```
3.4.3 Customizing Tooltip style
```{r}
tooltip_css<-"background-color:white;font-style:bold;color:black;"
p<-ggplot(data=exam_data,
          aes(x=MATHS))+
  geom_dotplot_interactive(
    aes(tooltip=ID),
    stackgroups=TRUE,
    binwidth=1,
    method="histodot")+
  scale_y_continuous(NULL,breaks = NULL
  )
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618,
  options=list(
    opts_tooltip(
      css=tooltip_css)
    )
  )
  
```
3.4.4 Displaying statistics on tooltip
```{r}
tooltip<-function(y,ymax,accuracy=.01){
  mean<-scales::number(y,accuracy=accuracy)
  sem<-scales::number(ymax-y,accuracy = accuracy)
  paste("Mean maths scores:",mean,"+/-",sem)
}
gg_point<-ggplot(data=exam_data,aes(x=RACE),
)+
  stat_summary(aes(y=MATHS,
          tooltip=after_stat(tooltip(y,ymax))),
          fun.data = "mean_se",
          geom=GeomInteractiveCol,
          fill="light blue"
  )+
  stat_summary(aes(y=MATHS),
               fun.data = mean_se,
               geom="errorbar",width=0.2,size=0.2)
girafe(ggobj=gg_point,
        width_svg=8,
        height_svg=8*0.618
        )

```
3.4.5 Hover effect with data_id aesthetic
```{r}
p<-ggplot(data=exam_data,
          aes(x=MATHS))+
            geom_dotplot_interactive(
              aes(data_id=CLASS),
              stackgroups=TRUE,
              binwidth=1,
              method="histodot")+
            scale_y_continuous(NULL,breaks=NULL)
girafe(
  ggobj=p,
  width_svg=6,
  height_svg=6*0.618
  )

```

3.4.6 Styling hover effect
```{r}
p<-ggplot(data=exam_data,
          aes(x=MATHS))+
  geom_dotplot_interactive(
    aes(data_id=CLASS),
    stackgroups=TRUE,
    binwidth=1,
    method="histodot")+
  scale_y_continuous(NULL,breaks=NULL)
girafe(
  ggobj=p,
  width_svg = 6,
  height_svg = 6*0.618,
  options=list(
    opts_hover(css="fill:#202020;"),
    opts_hover_inv(css="opacity:0.2;")
                         ))
                       
                     
    
```
3.4.7 Combining tooltip and hover effect
```{r}

```