---
title: "In-class Exercise 1"
author: "LIU YAN"
date: "15 April 2023"
editor: visual
---

# 1. Getting Started

## Install and Launching R Packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. if they are , then they will be lauched into R. - [tidyverse](https://www.tidyverse.org/): an opinionated collection of R packages designed for data science.

```{r}
pacman::p_load(tidyverse)
```

## Importing the Data

```{r}
exam_data<-read_csv("data/Exam_data.csv")
```

# 2. Excise

## 2.1 Working with Theme

```{r}
ggplot(data=exam_data, aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "lightblue", colour = "lightblue", 
                                    size = 0.5, linetype = "solid"),
    panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"), 
    panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "white"))
```

## 2.2 Designing Data-drive Graphics for Analysis I

```{r}
ggplot(data=exam_data, 
       aes(x=reorder(RACE,RACE,
                     function(x)-length(x)))) +
  geom_bar() +
  ylim(0,220) +
  geom_text(stat="count", 
      aes(label=paste0(..count.., ", ", 
      round(..count../sum(..count..)*100, 1), "%")),
      vjust=-1) +
  xlab("Race") +
  ylab("No. of\nPupils") +
  theme(axis.title.y=element_text(angle = 0))
```

## 2.3 Designing Data-drive Graphics for Analysis I

```{r}
exam_data %>%
  mutate(RACE = fct_infreq(RACE)) %>%
  ggplot(aes(x = RACE)) + 
  geom_bar()+
  ylim(0,220) +
  geom_text(stat="count", 
      aes(label=paste0(..count.., ", ", 
      round(..count../sum(..count..)*100,
            1), "%")),
      vjust=-1) +
  xlab("Race") +
  ylab("No. of\nPupils") +
  theme(axis.title.y=element_text(angle = 0))
```

## 2.4 Designing Data-drive Graphics for Analysis III

```{r}
d <- exam_data   
d_bg <- d[, -3]  

ggplot(d, aes(x = ENGLISH, fill = GENDER)) +
  geom_histogram(data = d_bg, fill = "grey", alpha = .5) +
  geom_histogram(colour = "black") +
  facet_wrap(~ GENDER) +
  guides(fill = FALSE) +  
  theme_bw()

```

## 2.5 Designing Data-drive Graphics for Analysis IV

```{r}
ggplot(data=exam_data, 
       aes(x=MATHS, y=ENGLISH)) +
  geom_point() +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  geom_hline(yintercept=50,
             linetype="dashed",
             color="grey60",
             size=1) + 
  geom_vline(xintercept=50, 
             linetype="dashed",
             color="grey60",
             size=1)
```

